About {\bf llvm2aa}

### 2011-04-27 v2/Ahir-asm/devel/docs/AaLRM.tex
reorganization

\newcommand{\Aa}{{\bf Aa}~}

@@

\section{Pointer de-reference expressions}

@@

\section{Address-of expressions}

@@

\section{Unary expressions}

@@

\section{Binary expressions}

@@

\section{Ternary expressions}

@@

>>>>

\newcommand{\Aa}{{\bf Aa}~}
\newcommand{\vC}{{\bf vC}~}

@@

\subsection{Pointer de-reference expressions}

@@

\subsection{Address-of expressions}

@@

\subsection{Unary expressions}

@@

\subsection{Binary expressions}

@@

\subsection{Ternary expressions}

@@

\section{Memory Spaces}

An \Aa program can contain declarations to storage objects.  These
storage objects need to be grouped into memory spaces using
the following rule: two storage objects are put in the same
memory space if there is a value in the program which
is infected by (determined by) the values of pointers to the two objects.

To understand this concept, consider the following
example
\begin{verbatim}
$module [foo] 
  $in (a: $uint<32>)
  $out (b: $uint<32>) 
$is
{
   $storage u: $uint<32>
   $storage v: $uint<32>
   $storage w: $uint<32>

   w := a
   ptr := @(u)
   ->(ptr) := w
   ptr := @(v)
   ->ptr := u
   b := u
}
\end{verbatim}
Now consider the object {\bf ptr}, which can hold a value which
is a pointer to either {\bf u} or a pointer to {\bf v}.  Thus,
{\bf u} and {\bf v} must belong to the same memory space.
However, {\bf w} can sit in a memory space by itself.  The
store accesses implied by the $->(ptr)$ will point to the
memory space which contains {\bf u} and {\bf v}.

What should we do in the following case?
\begin{verbatim}
$module [foo]
   $in (a : $uint<32>)
   $out (b: $uint<32>)
$is
{
   $storage u: $uint<32> 
   $storage ptr: $uint<32>
   ptr := ($bitcast ($pointer<$uint<32> >) a)
   tmp := ->(ptr)
   ptr := @(u)
   ->(ptr) := tmp
   b := u
}
\end{verbatim}
In this case, {\bf ptr} takes the value of $a$,
and as a pointer, it must point to 
an {\bf external} object.  Since ptr can 
point to this external object and also to {\bf u},
the external object and {\bf u} must be kept in
the same memory space.  In this case, the external
object needs
to be made internal, and moved {\bf inside} the system.  
This is done by the {\bf AaLinkExtMem} utility
which is described in a separate document.

### 2011-04-27 v2/llvmbc2Aa/docs/llvm2aa.tex
reorganization

\title{AhirV2: from algorithms to hardware \\ A tutorial introduction}
\author{Madhav Desai \\ Department of Electrical Engineering \\ Indian Institute of Technology \\
	Mumbai 400076 India}

@@

We describe the essential features
of the second version of the AhirV2 toolset
developed at IIT-Bombay.

\section{What is AhirV2?}

Perhaps the simplest way to understand AhirV2 is
through an example.

Suppose we start with a simple program.
\begin{verbatim}
int add(int a, int b)
{
        int c = (a+b);
        return(c);
}
\end{verbatim}
This program describes an algorithm.  We normally
compile this program and convert it to a machine-level
program which is executed on a processor.

What if we wished to convert this program to
a circuit?  
Such a circuit would have inputs 
$a,b$ and an output $c$ (together with some
interface hand-shake signals).  When activated,
the circuit should read its inputs and compute 
a response which is the same as would be expected
if the program were executed on a computer.

AhirV2 consists of a set of tools which takes
a C/C++ program and converts this collection to
a hardware circuit (described in VHDL).   This
conversion is done in two steps:
\begin{itemize}
\item The high-level program is compiled
to an interemediate assembly form.  AhirV2
introduces an intermediate assembly language
\Aa  which can serve as a target for sequential
programming languages (such as C/C++) as
well as for parallel programming languages.
\item From the \Aa description, a virtual
circuit (described in a virtual circuit
description language \vC) is generated.
The chief optimization carried out at this
step is that decomposing the system memory
into disjoint segments based on usage
analysis (this considerably improves the
available memory bandwidth and reduces
system cost).
\item From the \vC description, a
VHDL description of the system is generated.
The chief optimization carried out at this
stage is resource sharing.  The \vC description
is analyzed to identify operations which
cannot be concurrently active and this information
is used to reduce the hardware required.
\item 
The VHDL description produced from 
\vC is in terms of a library
of VHDL design units which has been developed
as part of the AhirV2  effort.  This library
consists of control-flow elements, data-path
elements and memory elements.
\item 
For the C/C++ to \Aa translation, the
AhirV2 toolset uses the front-end C/C++ compiler
infrastructure developed by the LLVM project (www.llvm.org), which
compiles the C/C++ program to LLVM byte-code
(which is a processor independent representation
of the compiled program).  The AhirV2 toolset
has a utility which translates the LLVM byte-code
to an \Aa description.
\end{itemize}
In Section \ref{sec:Example}, 
we illustrate the use of these tools on the
simple example shown above.

The earlier version of the AHIR flow has been
used to convert non-trivial programs to hardware.
The resulting circuits are upto two orders of
magnitude more energy-efficient than
a processor \cite{ref:dsd2010}.  The AhirV2
toolset incorporates several optimizations which
make the resulting circuits more 
competitive.

\section{The tools}

We assume that you have access to either {\bf llvm-gcc}
or {\bf clang} as the front-end compiler which generates
LLVM byte-code from C/C++.  The current AhirV2 toolset
is consistent with llvm 2.8 and clang 2.8.

The other tools in the chain are described below.

\subsection{{\bf llvm2aa}}

This tool takes LLVM byte code and converts it into an
\Aa file.
\begin{verbatim}
llvm2aa [-modules=<listfile>] [-storageinit] bytecode.o > bytecode.aa
\end{verbatim}
The generated \Aa code is sent to {\bf stdout} and all informational
messages are sent to {\bf stderr}.  On success, the tool returns 0.

>>>>

\title{About {\bf llvm2aa}}
\author{Madhav Desai \\ Department of Electrical Engineering \\ Indian Institute of Technology \\
	Mumbai 400076 India}

@@

{\bf llvm2aa}  is a tool which reads in LLVM byte-code (see http://www.llvm.org for
details about LLVM) and produces \Aa code which can then be further used
to produce VHDL using the AhirV2 tool chain developed at IIT Bombay.

\section{Synopsys}

The typical usage of the tool is 
\begin{verbatim}
llvm2aa [-modules=<listfile>] [-storageinit] bytecode.o > bytecode.aa
\end{verbatim}
The generated \Aa code is sent to {\bf stdout} and all informational
messages are sent to {\bf stderr}.  On success, the tool returns 0.
